---
title: "Titanic - Machine Learning from Disaster"
author: "Filip Wilhelm Sjöstrand"
date: "`r Sys.Date()`"
output: html_document
---



# Introduction

We are tasked with creating a prediction model on “what sorts of people were more likely to survive"? We shall create a logistic regression model with `Survived` as response variable. With this model, we will apply a test data set to see its predictions.

# Tidying

```{r, echo=F, message=F}
# Libraries ----------
library(dplyr)
library(ggcorrplot)
library(ggplot2)
library(tidyr)
```

```{r echo=F}
# Getting data ----------
train <- data.frame(read.csv('Data/train.csv',header=T, na.strings=c("")))
test <- data.frame(read.csv('Data/test.csv',header=T, na.strings=c("")))
```

We observe large missing values for `Cabin`, thus being dropped. Other insignificant values dropped are `PassengerId` since it is only an index, `Name`/`Ticket` since it is difficult to categorize, and the very few NAs in `Embarked`.

```{r}
# NA analysis ---------
train %>% visdat::vis_miss()
test %>% visdat::vis_miss()

train <- train %>%
    select(-PassengerId, -Cabin, -Name, -Ticket) %>%
    filter(!is.na(Embarked))

test <- test %>%
    select(-Cabin, -Name, -Ticket) %>%
    filter(!is.na(Embarked))
```

Given `Age`'s approximately normal distribution, I am going to assume no systematic missing NA and replace missing values with the average.

```{r}
train %>% ggplot() + geom_histogram(aes(Age))

train <- train %>% 
    mutate(Age = replace_na(Age, mean(train$Age,na.rm=T)))

test <- test %>% 
    mutate(Age = replace_na(Age, mean(test$Age,na.rm=T)))
```

We need to ensure every data type is correct. We can observe that `Survived`, `Pclass`, `Sex`, and `Embarked` are not factors—hence transformed.

```{r}
sapply(train, is.factor)

train <- train %>% 
    mutate(across(c(Survived, Pclass, Sex, Embarked), as.factor))

test <- test %>% 
    mutate(across(c(Pclass, Sex, Embarked), as.factor))
```

A last check on the summary values indicates nothing out of the ordinary.

```{r}
summary(train)
```


# Model Fitting

From first including every predictor, we observed that `Parch`, `Fare`, and `Embarked` were insignificant.

```{r}
model <- glm(
    Survived ~ Pclass + Sex + SibSp + Age,
    family = binomial(link = 'logit'),
    data = train
    )

summary(model)
```

## Diagnostics

Relatively high McFadden’s $R^2$

```{r}
pscl::pR2(model)["McFadden"]
```

No high VIF

```{r}
car::vif(model)
```

# Prediction

```{r}
preds <- predict(model, test, type="response")

test$prob <-  preds

result <- test %>%
    mutate(Survived = case_when(prob < .5 ~ 0, prob >= .5 ~ 1)) %>% 
    mutate(Survived = as.factor(Survived)) %>% 
    select(PassengerId, Survived)

path = "/Users/filipsjostrand/Documents/GitHub/Titanic/Data/Datasubmission.csv"
write.csv(result, path, row.names=FALSE)
```


